---
title: "Intro_To_R_Interactive"
author: "Dorcas Washington"
output: learnr::tutorial
runtime: shiny_prerendered
---


```{r setup, include=FALSE}
library(learnr)
```

## Workshop Summary and Contact Information

**Summary:** R is a free and powerful programming language that is commonly used by researchers in both qualitative and quantitative disciplines. R provides a near comprehensive, and still expanding set of research and data analysis tools. This workshop provides a gradual introduction to the basics of programming with R using R Studio. As a participant in this workshop you will accomplish the following: explore the R Studio interface, perform basic data manipulation, learn how to install and run packages, conduct simple data analyses, and produce basic data visualizations. The focus of this workshop will be hands-on exercises to provide a deeper and more effective understanding of R. No programming experience is required, and beginners are encouraged to attend.

**Contact:**   
Email: washids@uc.edu

Virtual Location: https://ucincinnati.webex.com/meet/washids

Physical Location : 240 Braunstein Hall (Geology-Math-Physics Library)

Research & Data Services Website: https://libraries.uc.edu/digital-scholarship/data-services.html

GitHub: dorcasmarie

This code below computes the answer to one plus one, change it so it computes two plus two:

```{r addition, exercise=TRUE}
1 + 1
```

R has feature called variables that can take on assignments. Below is an example. 

##### **Object and Assignments**
```{r assignment, exercise = FALSE}
x <- 1+2

y <- x +1
```


```{r letter-xy, echo=FALSE}
question("From above, what is the value of x and y?",
  answer("x = 3, y = 3"),
  answer("x = 3, y = 5"),
  answer("x = 3, y = 4", correct = TRUE),
  answer("x = 3, y = 2")
)
```

##### **Naming Conventions**
Keep names short, sweet, and relevent so anyone can follow your script!  
  
**Common Naming Conventions** 
Snake Case: "i_use_snake_case"   
Camel Case  "iUseCamelCase"  
Periods "i.use.periods"  
Combinations "iCant_decide.which_toUse"   
  
**Naming Hints**  
R is case sensative, so x != X  
Objects must start with a letter  
Use only numbers, letters, and _

##### **Understanding Functions**
Function formating: function_name(arg1 = val1, arg2 = val2)  
**HINT:** if you can't remember exactly what function you are looking we can use Tab    
Now that you found the function, How do you use it?  
**HINT:** Use the ? to find R documentation on an object  

```{r, eval = FALSE}
?seq()
```
##### <span style = "color:blue">**TASK 1A**</span>  
How would you create a sequence of numbers from 1 to 5 already shown below to 1 to 10 using the seq() function? 

```{r, sequence1, exercise = TRUE}
seq(1,5)
```

How would you create a sequence of numbers from 1 to 10 by 0.5 and save it to an object called Seq? Hint: review the documentation

```{r, sequence2, exercise = TRUE}
seq(1,10)

```

### **Working with Data Tables**

##### **Installing R package from CRAN Repository**
To access specific packages, that are not on base R, you must install that package before you can you use it. However, you only have to install a package once (per computer).  

Once the package is installed you have to run or open that package using the library function. This needs to be done everytime you open R Studio
```{r, eval=FALSE}
install.packages("tidyverse")
library(tidyverse)
```


##### **Quibbles with Tibbles**
Stolen from a tweet from Hadley Wickham  
**Tibbles** are modified versions of data.frames which make life in R a little easier.  
The tibble package and function is already installed under the tidyverse.  

Here is an example of a tibble with 4 variables with 10 values using diffent methods 

```{r, eval = FALSE}
vignette("tibble")

df_exp <- tibble::tibble(
  a = seq(1,10),
  b = rnorm(10),
  c = c(11:20),
  d = c(0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9,1.0)
)
```

##### <span style="color:blue">**TASK 2A**</span>  
What code would create a new tibble called "df" with two variables ("A" and "B") each with 20 different sets of observations?  

```{r, tibble, exercise = TRUE}

```

##### **Adding a column with mutate**
We will add a new column ("C"") to our table from <span style="color:blue">**TASK 2A**</span> using the mutate function from the dplyr package.    

We also will need to use the pipe (%>%) function which allows us to continue code, similar to an "and do this". 
**Hint** Using ctrl + shift + m is the shortcut key for pipe (%>%).  

```{r, eval =FALSE}
df <- df %>%
  dplyr::mutate(C = A/B)
```

##### **Exploring a data table**
Using brackets we can explore the values within certain positions of the table.  
For example, we want to see all rows in column 2 or the first 5 rows in each coloumn.

```{r, eval=FALSE}
df[,2]
df[5,]
```

##### <span style="color:blue">**TASK 2B**</span>  


```{r question_df, echo=FALSE}
question("What code would return the value in row 10 of column 3?",
  answer("df[3, 10]", message = "Close, but the syntax is df[row#(s), column#(s)]"),
  answer("df[10,,3]", message = "Close, but if you look there's a syntax error"),
  answer("df[10, 3]", correct=TRUE), random_answer_order=TRUE, allow_retry=TRUE


)
```
##### <span style="color:blue">**TASK 2C**</span>

```{r, question_more, echo=FALSE}
question("Based on this idea, What code would return the 3rd through 5th rows for the 2nd and 3rd columns? (Select all that apply)", 
  answer("df[3:5, 2:3]", correct = TRUE), 
  answer("df[c(3, 4, 5), c(2, 3)]", correct = TRUE), 
  answer("df[3:5, c(2, 3)]", correct = TRUE), random_answer_order=TRUE, allow_retry=TRUE)


```

##### **Saving a data table**
Now that we have created our data set we want to save it as a CSV file.  

```{r, eval=FALSE}
write.csv(df, "my_table.csv")
```


### **Exploring and Summarizing Data**

##### **Opening CSV Files**
Using the read.csv function we can open the CSV table data we created from before.    

```{r, eval=FALSE}
df <- read.csv("my_table.csv")
```

##### **Adding a new column to a data set**
We can add a new column to or existing data set by using the dollar sign ($) symbol with the name we want for our new column after symbol.  We want to add some non-numeric data to our table, so let use the letters a-t or letters that correspond to the 1st through 20th letter.

```{r, echo=TRUE, letters, exercise = TRUE}
df$Name <- letters[]

```

##### **Viewing a data set**
We can view our data in multiple ways but typically this is done by clicking the data set in the top left corner under the environment tab or using the View() function.  

Sometimes it is also important to view all of the columns or variables in a data set, this can be done using the names() function.  

Examining the structure of the data set is extremely important especially when we need to run analyses or visualize your data.  

**Hint** always view your data table and look at the structure of your data before working with a new data set.  

```{r, echo = TRUE, views, exercise = TRUE}
View(df)
names(df)
str(df)
```

##### **Summarizing Data and Descriptive Statistics**

Another very helpful function is summary(), which provides the basic overview of your data including min, max, quartiles, median, etc.   

If further statistical information is needed, use the stat.desc() function from the pastecs package as an efficient method to explore basic (similar to summary), descriptive statistical, and normal distribution information. Use ?stat.desc for more information.  

```{r, echo = TRUE, summary, exercise = TRUE}
summary(df)

#install.packages("pastecs")
library(pastecs)
?stat.desc()
stat.desc(df)
```
